SNP Analysis

### 2018-07-25 SNP_Prediction_Analysis.Rmd
generate new plots

ggplot(qcdata, aes(SCA.Ratio)) + geom_histogram() + 
  facet_wrap(~Status, scales='free_y', ncol=1) + labs(x='SCA Ratio', title='SCA Ratio, NP vs P')

ggplot(subset(qcdata, SampleType == 'BE'), aes(Purity, SCA.Ratio, color=Status)) + geom_point() + 
  geom_vline(xintercept = 0.95, color='red') + geom_hline(yintercept = 0.05, color='red') + 
  labs(title='Barretts only')

ggplot(subset(qcdata, SampleType == 'BE'), aes(Ploidy, SCA.Ratio, color=Status)) + geom_point() + 
  geom_hline(yintercept = 0.05, color='red') + labs(title='Barretts only')

@@

@@

```{r echo=F}
pred.qc.ex = subset(pred.qc, !exclude)
cuts = seq(0,1,0.1)

@@

@@

```{r regression, echo=F, message=F, eval=T}
## ----- All ----- ##
cache.dir = '~/Data/Ellie/Analysis/SNP'
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
 message(paste("loading", file))
 load(file, verbose=T)
} else {
  stop("Run 'snp-regression.R' first")
}
#save(arrayDf,labels,patient.info, file=paste(cache.dir,'snp-models.Rdata',sep='/'))

do.call(grid.arrange, c(plots))

model.performance(performance.at.1se, coefs)$plot + labs(title='SNP CV models')

@@

file = paste(cache.dir, 'loo.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  select.alpha = 1
  secf = coefs[[select.alpha]]
  a = select.alpha

  performance.at.1se = c(); plots = list(); fits = list(); nzcoefs = list()
  # Remove each patient (LOO)
  for (pt in unique(slabels$PatientID)) {
    print(pt)
    samples = subset(slabels, PatientID != pt)$Samplename

    train.rows = which(rownames(df) %in% samples)
    training = df[train.rows,]
    test = as.matrix(df[-train.rows,])
    if ( nrow(test) == ncol(df) ) test = t(test)

    # Predict function giving me difficulty when I have only a single sample, this ensures the dimensions are the same
    sparsed_test_data <- Matrix(data=0, nrow=nrow(test),  ncol=ncol(training),
                                dimnames=list(rownames(test),colnames(training)), sparse=T)
    for(i in colnames(arrayDf)) sparsed_test_data[,i] = test[,i]

    # Fit generated on all samples, including HGD
    fitLOO <- glmnet(training, status[train.rows], alpha=a, family='binomial', nlambda=nl, standardize=F) # all patients
    l = fitLOO$lambda

    cv = crossvalidate.by.patient(x=training, y=status[train.rows], lambda=l, a=a, nfolds=folds, splits=splits,
                                  pts=subset(sets, Samplename %in% samples), fit=fitLOO, standardize=F)

    plots[[pt]] = arrangeGrob(cv$plot+ggtitle('Classification'), cv$deviance.plot+ggtitle('Binomial Deviance'), top=pt, ncol=2)
    fits[[pt]] = cv

    if ( length(cv$lambda.1se) > 0 ) {
      performance.at.1se = c(performance.at.1se, subset(cv$lambdas, lambda == cv$lambda.1se)$mean)

      coef.1se = coef(fitLOO, cv$lambda.1se)[rownames(secf),]

      nzcoefs[[pt]] = as.data.frame(non.zero.coef(fitLOO, cv$lambda.1se))

      #coefs[[pt]] = coef(fitLOO, cv$lambda.1se)[rownames(secf),]
      #coef.stability(coef.1se, cv$non.zero.cf)

      logit <- function(p){log(p/(1-p))}
      inverse.logit <- function(or){1/(1 + exp(-or))}

      pm = predict(fitLOO, newx=sparsed_test_data, s=cv$lambda.1se, type='response')
      or = predict(fitLOO, newx=sparsed_test_data, s=cv$lambda.1se, type='link')

      slabels[which(slabels$Samplename %in% rownames(pm)),'Prediction'] = pm[,1]
      slabels[which(slabels$Samplename %in% rownames(pm)),'RR'] = or[,1]

    } else {
      warning(paste(pt, "did not have a 1se"))
    }
  }
  save(plots, performance.at.1se, nzcoefs, fits, slabels, file=file)
}

snp.roc = pROC::roc(Status~Prediction,slabels, auc=T, ci=T, of='thresholds')
roc.plot(snp.roc, title="SNP LOO Model ROC")

>>>>

ggplot(subset(qcdata, SampleType == 'BE'), aes(round(Ploidy,2), Purity, color=Status)) + geom_point() + facet_grid(~Status) + 
  labs(title='BE only, purity vs ploidy')

ggplot(qcdata, aes(SCA.Ratio, color=Status, fill=Status)) + geom_density(alpha=0.5) + labs(x='SCA Ratio', title='SCA Ratio, NP vs P')

ggplot(subset(qcdata, SampleType == 'BE'), aes(Purity, SCA.Ratio, color=Status)) + facet_grid(~Status) +
  geom_point() + geom_vline(xintercept = 0.5, color='darkgreen') + geom_hline(yintercept = 0.05, color='red') + 
  labs(title='Barretts only')

ggplot(subset(qcdata, SampleType == 'BE'), aes(Ploidy, SCA.Ratio, color=Status)) + facet_grid(~Status) + geom_point() + 
  geom_hline(yintercept = 0.03, color='red') + labs(title='Barretts only')

@@

subset(pred.qc, `mean(Ploidy)` < 1.99)

subset(pred.qc, lowsca)

@@

```{r echo=F}
pred.qc.ex = subset(pred.qc, !exclude & !wgd & !lowsca)
cuts = seq(0,1,0.1)

@@

roc = pROC::roc(Status ~ Prob, data=subset(pred.qc, !exclude & !wgd), auc=T, ci=T, of='thresholds')
roc.plot(roc) + labs(title='Excluded HGD & WGD NP patients')

@@

```{r regression, echo=F, message=F, eval=F}
## ----- All ----- ##
cache.dir = '~/Data/Ellie/Analysis/SNP'
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
 message(paste("loading", file))
 load(file, verbose=T)
 model.performance(performance.at.1se, coefs)$plot + labs(title='SNP CV models')
} else {
  warning("Run 'snp-regression.R' first")
}
#save(arrayDf,labels,patient.info, file=paste(cache.dir,'snp-models.Rdata',sep='/'))

#do.call(grid.arrange, c(plots))
```

@@

file = paste(cache.dir, 'loo.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  stop("Run snp-regression.R first")
}

slabels
snp.roc = pROC::roc(Status~Prediction,slabels, auc=T, ci=T, of='thresholds')
roc.plot(snp.roc, title="SNP LOO Model ROC")
